---
title: "Overview"
---

```{r, eval = TRUE, echo = FALSE}
source("R/utils.R")
```

R-multiverse provides:

1. A home for packages that fall outside the scope of other repositories such as CRAN and Bioconductor.
2. Direct and timely distribution of package releases.
3. Assurance of package quality for production scenarios.

## R-multiverse repositories

R-multiverse hosts two R package repositories:

1. [Community](community.md): an open, inclusive, community-curated repository
of R package releases pulled directly from their maintainers via GitHub and GitLab.
2. [Production](production.qmd): a quarterly snapshot from [Community](community.md)
of releases which pass transparent automated checks.

## Installing packages

Each R-multiverse repository has a URL for the `repos` argument of `install.packages()`.
For example, to install [`polars`](https://pola-rs.github.io/r-polars/) from [Community](community.md):
^[
Caution: when setting two repositories in the `repos` argument,
the same package may be in both.
`install.packages()` chooses the most current version,
which could be in either repository.
]

```r
install.packages(
  "polars",
  repos = c("https://community.r-multiverse.org", getOption("repos"))
)
```

[Production](production.qmd) is deployed in periodic snapshots throughout the year.
The current Production snapshot was deployed on `r snapshot()$snapshot`.
It was tested with the base R `r snapshot()$r` and [CRAN](https://cran.r-project.org/) packages from the [dependency freeze](#staging) on `r snapshot()$dependency_freeze`.
Please use the `r snapshot()$dependency_freeze` CRAN snapshot from [Posit Public Package Manager (p3m)](https://packagemanager.posit.co) to install dependencies:

```{r, eval = TRUE, echo = FALSE, results = "asis"}
text <- c(
  "```r",
  "install.packages(",
  "  \"polars\",",
  "  repos = c(",
  "    \"https://production.r-multiverse.org\",",
  sprintf("    \"https://packagemanager.posit.co/cran/%s\"", snapshot()$dependency_freeze),
  "  )",
  ")",
  "```"
)
cat(text, sep = "\n")
```

## Other repositories

R-multiverse does not replace [CRAN](https://cran.r-project.org/).
In fact, most R-multiverse packages depend on [CRAN](https://cran.r-project.org/) packages.
In production environments,
[Production](production.qmd) snapshots should be deployed alongside
the [CRAN](https://cran.r-project.org/) package dependencies
and release version of base R from the same day that the R-multiverse snapshot was created.

## Infrastructure

All [R-multiverse source code](https://github.com/r-multiverse) is public and open-source.
The code is licensed under [MIT](https://github.com/r-multiverse/multiverse.internals/blob/main/LICENSE.md), and the [logo](https://github.com/r-multiverse/r-multiverse.github.io/tree/main/logo) has a [Creative Commons Attribution-ShareAlike 4.0 International license](https://github.com/r-multiverse/r-multiverse.github.io/blob/main/logo/LICENSE.md). 
R-multiverse leverages [R-universe](https://r-universe.dev/), [GitHub Actions](https://github.com/features/actions), and [Cloudflare](https://www.cloudflare.com/).
[Cloudflare](https://www.cloudflare.com/) provides the `r-multiverse.org` domain and hosts [Production](production.qmd) snapshots in an [R2 bucket](https://www.cloudflare.com/developer-platform/products/r2/).

## How you can help

* Spread the word: share R-multiverse with your colleagues.
* [Contribute your packages](contributors.md) to R-multiverse.
* Volunteer to become a [moderator](moderators.md).

## Support

R-multiverse originated from the R Consortium [Repositories Working Group](https://github.com/RConsortium/r-repositories-wg), and it was supported by [R Consortium](https://r-consortium.org) ISC grant 24-ISC-1-05.
